### 2021-09-15 thesis/tex/oscillations.tex
update

\section{Neutrino Mixing}\label{ch:oscillation}
Japan 1998. The Super-Kamiokande experiment detects a $\nm$ deficit that could not be explained by any other mechanism
than the conversion of muon neutrinos to electron neutrinos. This was the first conclusive evidence of a completely new physics phenomenon: neutrino flavor
oscillations. An implication of the observation of oscillations is neutrino mass, 
which is identically zero according to the Standard Model. Thus, we are required to extend the Standard Model to incorporate neutrino 
masses before we discuss the neutrino oscillations themselves.

\subsection{Mass Generation}
In the Standard Model, fermion masses are generated by the Higgs mechanism through Yukawa couplings with the fermion's right and left-handed components.
All neutrinos are left-handed and all antineutrinos are right-handed. Thus, they do not pick up a Yukawa coupling, 
and will not undergo the Higgs mechanism, which leaves them massless.
Additionally, all terms of a Lagrangian that we construct must respect the gauge invariance. This removes the possibility of the neutrino mass to be 
generated at loop level because any such attempt will violate the total lepton number by two units.
In order to keep our theory renormalizable, we have one option left if we want to generate neutrino masses: introducing right-handed neutrinos and left-handed antineutrinos.

@@

We extend the Standard Model by adding a right-handed component of this field with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ to the Higgs-lepton Yukawa Lagrangian 
\begin{align}
    \mathcal{L}_{H}=-\left( \frac{v + H}{\sqrt{2}} \right) \left[\ell_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \nu_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]\,.
\end{align}
Now, the Yukawa coupling matrices $Y^{\prime \ell}$ and $Y^{\prime \nu}$ are non-diagonal, with the former resulting in the charged lepton masses being non-definite. 
This can be remedied by diagonalizing $Y^{\prime \ell}$ with a unitary matrix $V^\ell$ as
\begin{align}
    V_{\alpha k L}^{\nu \dagger} Y^{\prime \nu}_{\alpha \beta} V_{\beta j R}^{\nu}=Y^{\nu}_{kj} \,.
\end{align}
Similarly, we diagonalize and $Y^{\prime \nu}$ with a unitary matrix $V^\nu$. We note that this will not yet generate neutrino masses, 
since no flavor rotation can make massless fields massive.
Now we state a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. 
This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement for neutrino flavor eigenstates to have a definite mass. 
The flavor of a neutrino is then, by definition, the flavor of the associated charged lepton. 
This fact is denoted as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.

@@

Call $V^{\prime \nu \dagger}_{k \alpha}V^{\prime \ell}_{\alpha \alpha} = U^\dagger_{k \alpha}$. We will refer to the matrix $U$ built by the components $U_{k \alpha}$ as the
Pontecorvo-Maki-Nakagawa-Sakata (PMNS) matrix~\cite{pontecorvo1957,maki1962}. %TODO: source
We now have 
\begin{align}\label{eq:j_CC3} %TODO: put sums on all above?
    j^\rho_L &= 2 \sum_\alpha \sum_k U^\dagger_{\alpha k} \bar{\nu}_{k L} \gamma^\rho  \ell_{\alpha L}\,.
\end{align}

What form does this new matrix $U$ take? By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three mixing angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable by us. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure. We are now down to four degrees of freedom in the three dimensional case: three mixing angles
of the form $\sin{(\theta_{ij})}/\cos{(\theta_{ij})} = s_{ij}/c_{ij}$, and one phase of the form $e^{i\delta}$. The mixing angle $\theta_{ij}$ specifies the amount of rotation in the $i-j$ plane, allowing an $i$ eigenstate to mix with a $j$ eigenstate. The physical interpretation of
a mixing angle is then the degree to which the $\nu_i$ eigenstate mixes with the $\nu_j$ eigenstate. A non-zero mixing angle will then 
cause the flavor eigenstate $\nu_\alpha$ to be a superposition of the mass eigenstates that we allowed to mix. %TODO: nua = nu1 + nu2

We construct the PMNS matrix using the rotation matrixes from $SO(3)$ as
\begin{align}\label{PMNS_def}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12} \nonumber \\
      & = 
    \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix}\,.
\end{align}
The term $\delta_\text{CP}$ determines the degree of charge-parity violation. In this work, we will always assume CP invariance for simplicity by
setting $\delta_{CP} = \SI{0}{\degree}$. Thus,

@@

In this work, we will be using the best-fit values from~\cite{nufit} with the exception of the CP-violating phase $\delta_\text{CP}$. The $90\% $ CL values are
\begin{align}\label{eq:nufitparams}
    \theta_{12} = \SI{33.44}{\degree},\hspace{1em} \theta_{13} = \SI{8.57}{\degree},\hspace{1em} \theta_{23} = \SI{49.2}{\degree}, \hspace{1em} \delta_\text{CP} = 0\,.
\end{align}
Our neglect of the issue of CP invariance leaves the PMNS matrix real, with numerical values 
\begin{align}\label{eq:Uvalues}
    U = \begin{pmatrix}
        U_{e 1} & U_{e2} & U_{e3} \\
        U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
        U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
    \end{pmatrix} 
    = \begin{pmatrix}
        0.825 & 0.545 & 0.149 \\
        -0.455 & 0.485 & 0.746 \\
        0.334 & -0.684 & 0.649
    \end{pmatrix} \,.
\end{align}
In our study, we will only use neutrinos originating from conventional cosmic ray interactions in the atmosphere. 
In this so-called atmospheric neutrino flux, $\nm$ are the most abundant, while the $\nt$ flux is negligible. From Eq.~\eqref{eq:Uvalues}, w
e see that $U_{\mu\tau}$ is the largest $U_{\mu\beta}$ term,
suggesting that $\nm$ primarily mixes into $\nt$, at least in vacuum. It turns out that matter effects indeed preserves this ordering, making the $\nm \to \nt$ transition
the most abundant. This means we will be able to stringently constrain parameters relating to $\mu\tau$ mixing.

\subsection{Neutrino Propagation}
Since we now know how the neutrino mass and flavor eigenstates combine, and have an expression for the flavor interaction with 
the neutrino's charged lepton partner, we are now ready to study the flavor oscillations themselves.

Now, the Fourier expansion of the field operator $\bar{\nu}_{k L}$ in the current of Eq.~\ref{eq:j_CC3} contains creation operators $a^\dagger_{\nu_k}$
of massive neutrinos with mass $m_k$. This means that the summation over the mass index $k$ constructs a flavor neutrino, which interacts with the charged lepton field $\ell_{\a L}$.
In other words, the charged current generates a flavor neutrino $\na$, which is a superposition of the mass eigenstates 
$\nu_k$ with weights $U_{\a k}^\dagger$. In the ket-formalism, we express this as
\begin{align}\label{eq:osc_1}
    \ket{\na} = \sum_k U^\dagger_{\a k} \ket{\nu_k}\,.
\end{align}
It is the mass eigenstates $\ket{\nu_k}$ that are eigenstates of the Hamiltonian, with eigenvalues
\begin{align}\label{eq:disp}
    E_k = \sqrt{\vec{p}^2 + m_k^2}\,.
\end{align}
The solution to the time-dependent Schrödinger equation 
\begin{align}\label{eq:TDSE}
    i \dv{}{t} \ket{\nu_k(t)} = H_0\ket{\nu_k(t)}\,,
\end{align}
where $H_0$ is the vacuum Hamiltonian.

The solution to Eq.~\ref{eq:TDSE} gives ut the time evolution
in the form of plane wave solutions: 
\begin{align}
    \ket{\nu_k(t)} = e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Inserting the plane wave solution into Eq.~\ref{eq:osc_1}, we get 
\begin{align}\label{eq:nat}
    \ket{\na(t)} = \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Now we know how to evolve and combine the mass eigenstates to form a flavor eigenstate, but how about the reverse?We swap the index $k\to j$ in Eq.~\ref{eq:osc_1} and multiply by $U_{\a k}$:
\begin{align}\label{eq:osc_k}
    \sum_\a U_{\a k}\ket{\na} &= \sum_{\alpha, j} U_{\a k} U_{\a j}^\dagger \ket{\nu_j} \nonumber \\
                              &= \sum_{j} \delta_{kj} \ket{\nu_j} \nonumber \\
                              &= \ket{\nu_k}\,,
\end{align}
where we have used the unitarity of the leptonic mixing matrix. Eqs.~\ref{eq:osc_1} and \ref{eq:nat} yield 
\begin{align}
    \ket{\na(t)} &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k} \nonumber \\
                 &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\left(\sum_\b U_{\b k}\ket{\nb}\right) \nonumber \\
                 &= \sum_{k,\b} U^\dagger_{\a k}U_{\b k} e^{-iE_kt} \ket{\nb}\,.
\end{align}
The probability of the flavor transition $\na \to \nb$ at time $t$ is $\abs{\bra{\nb}\ket{\na (t)}}^2$:
\begin{align}
    P_{\na \to \nb}(t) = \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(E_k-E_j)t}\,.
\end{align}

We assume the neutrino masses $m_k$ to be extremely small compared to their associated energies $E_k$. 
Thus, $v\sim 1$, and $\abs{\vec{p}} \sim E$ making the energy-dispersion relation of Eq.~\ref{eq:disp} to first order:
\begin{align}\label{eq:ultra_rel}
    E_k &= \sqrt{\vec{p}^2 + m_k^2} \nonumber \\
        &= \vec{p}^2\sqrt{1 + \frac{m_k^2}{\vec{p}^2}} \nonumber \\
        &\approx E + \frac{m_k^2}{2E}
\end{align}
Hence, the exponential can be simplified, and simplifying the notation $P_{\na \to \nb}(t) \to P_{\ab}(t)$ we get 
\begin{align}
    P_{\ab}(t) = \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(m_k^2-m_j^2)t/2E}\,.
\end{align}
Now, our approximation $v\approx 1$ implies $x\approx t $, thus
\begin{align}
    P_{\ab}(x) &= \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(m_k^2-m_j^2)x/2E} \nonumber \\
                       &= \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E})\,,
\end{align}
where we in the last step have defined the \emph{mass-squared difference} $\dm_{kj} = m_k^2 - m_j^2$. Since the oscillation probability depends on this quantity
rather than the individual masses, it is impossible to measure the mass $m_k$ through neutrino oscillations.
Squaring the unitarity condition $\sum_k U_{\a k}U_{\b k}^\dagger = \delta_{\a \b}$ yields 
\begin{align}
    \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 = \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger]
\end{align}
Since we take the PMNS matrix to be real and thus Hermitian, we drop the complex transpose in the following calculation. We have
\begin{align}\label{eq:Pab}%TODO: shorten?
    P_{\ab} &= \sum_{k,j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &=  \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 + \sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} +\sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} + 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\exp(-i\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\cos(\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}
which is the probability of neutrino vacuum oscillations.
The calculation for antineutrinos yield the same result if one continues to assume the realness of the mixing matrix. 
In Fig.~\ref{fig:vac_osc}, we show the energy spectra of these oscillations in the low \si{\GeV} range for neutrinos that travel \SI{12000}{\km}\footnote{This baseline
is approximately the Earth diameter, and will be the maximum travel distance which we will use later on.}.
In the left-most panel, we see a supressed $\nm \to \ne$ transition. We can understand this behavior by studying Eq.~\ref{eq:Pab} and the 
numerical values of the mixing matrix in Eq.~\ref{eq:Uvalues}.
\si{\GeV} oscillations are driven by the $\dm[31]$ mass-squared difference. Hence, the most important mixing matrix 
elements for $P_{\a\b}$ are $U_{\a3}, U_{\a1}$ and $U_{\b3}, U_{\b1}$. The smallness of $U_{e3}$ then supresses the
amplitude of $\Pme$.
\begin{figure}[!h]
    \begin{center}
    \includegraphics[width=0.9\textwidth]{figures/vac_osc.pdf}
    \caption{$\nm$ vacuum oscillations from Eq.~\ref{eq:Pab}}\label{fig:vac_osc}
    \end{center}
\end{figure}

this form elucidates an important aspect of neutrino oscillations. The mixing matrix elements
determine the amplitude of the oscillations, while the mass-squared differences together with the ratio $L/E$ determine the 
frequency. So in vacuum, the mixing angles only influence the oscillation amplitude, while the energy, trajectory, and masses of
the neutrinos only influence the oscillation frequency.

@@

For neutral current, we replace the electron field $e(x)$ in Eq.~\ref{eq:H_fierz} by the more general fermion field $f(x)$, and the projection operator 
$(1-\gamma^5)$ with $(g_V^f - g_A^f\gamma^5)$. Again, the $\gamma^5$ will cause the spacial component of $p_f$ to disappear after integration, and the 
only difference between the average effective Hamiltonian for the neutral current is then the factor $g_V^f$:
\begin{align}
    V^f_{NC} = \sqrt{2}G_F N_A g_V^f\,.
\end{align}
Summing over the fermions, and assuming electrical neutrality and equal abundance of protons and neutrons, we have
\begin{align}
    V_{NC} &= \sum_{f \in {e,p,n}} V^f_{NC} \nonumber \\
           &= \sqrt{2}G_F N_A \sum_{f \in {e,p,n}} g_V^f \nonumber \\
           &= \sqrt{2}G_F N_A\left[ -\frac{1}{2}+2\sin^2{(\theta_W)} + \frac{1}{2}-2\sin^2{(\theta_W)} -\frac{1}{2} \right] \nonumber \\
           &= -\frac{1}  {\sqrt{2}} G_F N_e\,,
\end{align}
where the electrical neutrality condition allows us to simply sum the vectorial couplings together, cancelling the electron and proton contributions (and hence, also the $\theta_W$ dependence).

>>>>

\section{Neutrino Mixing}\label{ch:oscillation}
Japan 1998. The Super-Kamiokande experiment detects a $\nm$ deficit that could not be explained by any other mechanism
than the conversion of muon neutrinos to electron neutrinos~\cite{sk1998}. This was the first conclusive evidence of a completely new physics phenomenon: neutrino flavor
oscillations. An implication of the observation of oscillations is neutrino mass, 
which is identically zero according to the Standard Model. Thus, we are required to extend the Standard Model to incorporate neutrino 
masses before we discuss the neutrino oscillations themselves.

\subsection{Mass Generation}
In the Standard Model, fermion masses are generated by the Higgs mechanism through Yukawa couplings with the fermion's right and left-handed components.
All neutrinos are left-handed and all antineutrinos are right-handed~\cite{giunti}. Thus, they do not pick up a Yukawa coupling, 
and will not undergo the Higgs mechanism, which leaves them massless.
Additionally, all terms of a Lagrangian that we construct must respect the gauge invariance. This removes the possibility of the neutrino mass to be 
generated at loop level because any such attempt will violate the total lepton number by two units.
In order to keep our theory renormalizable, we have one option left if we want to generate neutrino masses: introducing right-handed neutrinos and left-handed antineutrinos.

@@

We extend the Standard Model by adding a right-handed component of this field with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ to the Higgs-lepton Yukawa Lagrangian 
\begin{align}
    \mathcal{L}_{H}=-\left( \frac{v + H}{\sqrt{2}} \right) \left[\ell_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \nu_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]\,,
\end{align}
where $v$ is the Higgs vacuum expectation value, $H$ the Higgs field, $\ell_{\alpha L}^\prime (\ell_{\beta R}^\prime)$ the left (right) handed charged lepton field with non-definite masses. $Y_{\alpha \beta}^{\prime \ell}$ are the charged lepton Yukawa couplings, and $\nu^\prime_{\alpha L}(\nu^\prime_{\beta R})$ are the left (right) handed neutrino fields .
Now, the Yukawa coupling matrices $Y^{\prime \ell}$ and $Y^{\prime \nu}$ are non-diagonal, with the former resulting in the charged lepton masses being non-definite. 
This can be remedied by diagonalizing $Y^{\prime \ell}$ with a unitary matrix $V^\ell$ as
\begin{align}
    V_{\alpha k L}^{\nu \dagger} Y^{\prime \nu}_{\alpha \beta} V_{\beta j R}^{\nu}=Y^{\nu}_{kj} \,.
\end{align}
Similarly, we diagonalize and $Y^{\prime \nu}$ with a unitary matrix $V^\nu$. We note that this will not yet generate neutrino masses, 
since no flavor rotation can make massless fields massive.
Now we state a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. 
This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement for neutrino flavor eigenstates to have a definite mass. 
The flavor of a neutrino is then, by definition, the flavor of the associated charged lepton. 
This fact is denoted as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.

@@

Call $V^{\prime \nu \dagger}_{k \alpha}V^{\prime \ell}_{\alpha \alpha} = U^\dagger_{k \alpha}$. We will refer to the matrix $U$ built by the components $U_{k \alpha}$ as the (neutrino) mixing matrix. It can also be referred to as the 
Pontecorvo-Maki-Nakagawa-Sakata (PMNS) matrix~\cite{pontecorvo1957,maki1962}.
We now have 
\begin{align}\label{eq:j_CC3} %TODO: put sums on all above?
    j^\rho_L &= 2 \sum_\alpha \sum_k U^\dagger_{\alpha k} \bar{\nu}_{k L} \gamma^\rho  \ell_{\alpha L}\,.
\end{align}

What form does this new matrix $U$ take? By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three mixing angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable by us. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure. We are now down to four degrees of freedom in the three dimensional case: three mixing angles
of the form $\sin{(\theta_{ij})}/\cos{(\theta_{ij})} = s_{ij}/c_{ij}$, and one phase of the form $e^{i\delta}$. The mixing angle $\theta_{ij}$ specifies the amount of rotation in the $i-j$ plane, allowing an $i$ eigenstate to mix with a $j$ eigenstate. The physical interpretation of
a mixing angle is then the degree to which the $\nu_i$ mass eigenstate mixes with the $\nu_j$ mass eigenstate. A non-zero mixing angle will then 
cause the flavor eigenstate $\nu_\alpha$ to be a superposition of the mass eigenstates that we allowed to mix. More on this in Section~\ref{sec:prop}.

We construct the PMNS matrix using the rotation matrices from $SO(3)$ as
\begin{align}\label{PMNS_def}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12} \nonumber \\
      & = 
    \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix}\,.
\end{align}
The term $\delta_\text{CP}$ determines the degree of charge-parity violation. In this work, we will always assume CP invariance for simplicity by
setting $\delta_{CP} = \SI{0}{\degree}$. Thus,

@@

In this work, we will be using the best-fit values from NuFit~\cite{nufit} with the exception of the CP-violating phase $\delta_\text{CP}$. The $90\% $ CL values are
\begin{align}\label{eq:nufitparams}
    \theta_{12} = \SI{33.44}{\degree},\hspace{1em} \theta_{13} = \SI{8.57}{\degree},\hspace{1em} \theta_{23} = \SI{49.2}{\degree}, \hspace{1em} \delta_\text{CP} = 0\,.
\end{align}
Our neglect of the issue of CP invariance leaves the PMNS matrix real, with numerical values 
\begin{align}\label{eq:Uvalues}
    U = \begin{pmatrix}
        U_{e 1} & U_{e2} & U_{e3} \\
        U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
        U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
    \end{pmatrix} 
    = \begin{pmatrix}
        0.825 & 0.545 & 0.149 \\
        -0.455 & 0.485 & 0.746 \\
        0.334 & -0.684 & 0.649
    \end{pmatrix} \,.
\end{align}
In our study, we will only use neutrinos originating from conventional cosmic ray interactions in the atmosphere. 
In this so-called atmospheric neutrino flux, $\nm$ are the most abundant, while the $\nt$ flux is negligible. From Eq.~\eqref{eq:Uvalues}, we see that $U_{\mu\tau}$ is the largest $U_{\mu\beta}$ term,
suggesting that $\nm$ primarily mixes into $\nt$, at least in vacuum. It turns out that matter effects indeed preserves this ordering, making the $\nm \to \nt$ transition
the most abundant.

\subsection{Neutrino Propagation}\label{sec:prop}
Since we now know how the neutrino mass and flavor eigenstates combine and have an expression for the flavor interaction with 
the neutrino's charged lepton partner, we are now ready to study the flavor oscillations themselves.

Now, the Fourier expansion of the field operator $\bar{\nu}_{k L}$ in the current of Eq.~\ref{eq:j_CC3} contains creation operators $a^\dagger_{\nu_k}$
of massive neutrinos with mass $m_k$~\cite{giunti}. This means that the summation over the mass index $k$ with the mixing matrix elements constructs a flavor neutrino, which interacts with the charged lepton field $\ell_{\a L}$.
In other words, the charged current generates a flavor neutrino $\na$, which is a superposition of the mass eigenstates 
$\nu_k$ with weights $U_{\a k}^\dagger$. In the ket-formalism, we express this as
\begin{align}\label{eq:osc_1}
    \ket{\na} = \sum_k U^\dagger_{\a k} \ket{\nu_k}\,.
\end{align}
It is the mass eigenstates $\ket{\nu_k}$ that are eigenstates of the Hamiltonian, with eigenvalues
\begin{align}\label{eq:disp}
    E_k = \sqrt{\vec{p}^2 + m_k^2}\,.
\end{align}
The solution to the time-dependent Schrödinger equation 
\begin{align}\label{eq:TDSE}
    i \dv{}{t} \ket{\nu_k(t)} = H_0\ket{\nu_k(t)}\,,
\end{align}
where $H_0$ is the vacuum Hamiltonian. 
The solution to Eq.~\ref{eq:TDSE} gives ut the time evolution
in the form of plane wave solutions: 
\begin{align}
    \ket{\nu_k(t)} = e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Inserting the plane wave solution into Eq.~\ref{eq:osc_1}, we get 
\begin{align}\label{eq:nat}
    \ket{\na(t)} = \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Now we know how to evolve and combine the mass eigenstates to form a flavor eigenstate, but how about the reverse? We swap the index $k\to j$ in Eq.~\ref{eq:osc_1} and multiply by $U_{\a k}$:
\begin{align}\label{eq:osc_k}
    \sum_\a U_{\a k}\ket{\na} &= \sum_{\alpha, j} U_{\a k} U_{\a j}^\dagger \ket{\nu_j} \nonumber \\
                              &= \sum_{j} \delta_{kj} \ket{\nu_j} \nonumber \\
                              &= \ket{\nu_k}\,,
\end{align}
where we have used the unitarity of the leptonic mixing matrix. Eqs.~\ref{eq:osc_1} and \ref{eq:nat} yield 
\begin{align}
    \ket{\na(t)} &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k} \nonumber \\
                 &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\left(\sum_\b U_{\b k}\ket{\nb}\right) \nonumber \\
                 &= \sum_{k,\b} U^\dagger_{\a k}U_{\b k} e^{-iE_kt} \ket{\nb}\,.
\end{align}
The probability of the flavor transition $\na \to \nb$ at time $t$ is $\abs{\bra{\nb}\ket{\na (t)}}^2$:
\begin{align}
    P_{\na \to \nb}(t) &= {\bra{\nb}\ket{\na (t)}}^2\sum_{k,j} \nonumber \\
    &= U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(E_k-E_j)t}\,.
\end{align}

We assume the neutrino masses $m_k$ to be extremely small compared to their associated energies $E_k$. 
Thus, $v\approx 1$, and $\abs{\vec{p}} \approx E$ making the energy-dispersion relation of Eq.~\ref{eq:disp} to first order:
\begin{align}\label{eq:ultra_rel}
    E_k &= \sqrt{\vec{p}^2 + m_k^2} \nonumber \\
        &= \vec{p}^2\sqrt{1 + \frac{m_k^2}{\vec{p}^2}} \nonumber \\
        &\approx E + \frac{m_k^2}{2E}
Hence, the exponential can be simplified, and simplifying the notation $P_{\na \to \nb}(t) \to P_{\ab}(t)$ we get 
    P_{\ab}(t) = \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(m_k^2-m_j^2)t/2E}\,.
Now, our approximation $v\approx 1$ implies $x\approx t $, thus
    P_{\ab}(x) &= \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(m_k^2-m_j^2)x/2E} \nonumber \\
                       &= \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E})\,,
where we in the last step have defined the \emph{mass-squared difference} $\dm_{kj} = m_k^2 - m_j^2$. Since the oscillation probability depends on this quantity
rather than the individual masses, it is impossible to measure the mass $m_k$ through neutrino oscillations. 
Squaring the unitarity condition $\sum_k U_{\a k}U_{\b k}^\dagger = \delta_{\a \b}$ yields 
    \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 = \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger]
Since we take the PMNS matrix to be real and thus Hermitian, we drop the complex transpose in the following calculation. We have
\begin{align}\label{eq:Pab}%TODO: shorten?
    P_{\ab} &= \sum_{k,j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &=  \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 + \sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} +\sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} + 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\exp(-i\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\cos(\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
which is the probability of neutrino vacuum oscillations.
The calculation for antineutrinos yield the same result if one continues to assume the realness of the mixing matrix. 
In Fig.~\ref{fig:vac_osc}, we show the energy spectra of these oscillations in the low \si{\GeV} range for neutrinos that travel \SI{12000}{\km}\footnote{This baseline
is approximately the Earth diameter, and will be the maximum travel distance which we will use later on.}.
In the left-most panel, we see a supressed $\nm \to \ne$ transition. We can understand this behavior by studying Eq.~\ref{eq:Pab} and the 
numerical values of the mixing matrix in Eq.~\ref{eq:Uvalues}.
\si{\GeV} oscillations are driven by the $\dm[31]$ mass-squared difference. Hence, the most important mixing matrix 
elements for $P_{\a\b}$ are $U_{\a3}, U_{\a1}$ and $U_{\b3}, U_{\b1}$. The smallness of $U_{e3}$ then supresses the
amplitude of $\Pme$.
\begin{figure}[!h]
    \begin{center}
    \includegraphics[width=0.9\textwidth]{figures/vac_osc.pdf}
    \caption{$\nm$ vacuum oscillations from Eq.~\ref{eq:Pab}
    for $\nm$ neutrinos travelling 12 000 km.}\label{fig:vac_osc}
    \end{center}
\end{figure}
This form elucidates an important aspect of neutrino oscillations. The mixing matrix elements
determine the amplitude of the oscillations, while the mass-squared differences together with the ratio $L/E$ determine the 
frequency. So in vacuum, the mixing angles only influence the oscillation amplitude, while the energy, trajectory, and masses of
the neutrinos only influence the oscillation frequency.

@@

For neutral current, we replace the electron field $e(x)$ in Eq.~\ref{eq:H_fierz} by the more general fermion field $f(x)$, and the chiral projection operator 
$(1-\gamma^5)$ with $(g_V^f - g_A^f\gamma^5)$. Again, the $\gamma^5$ will cause the spacial component of $p_f$ to disappear after integration, and the 
only difference between the average effective Hamiltonian for the neutral current is then the factor $g_V^f$:
\begin{align}
    V^f_{NC} = \sqrt{2}G_F N_A g_V^f\,.
\end{align}
Summing over the fermions, and assuming electrical neutrality and equal abundance of protons and neutrons, we have
\begin{align}
    V_{NC} &= \sum_{f \in {e,p,n}} V^f_{NC} \nonumber \\
           &= \sqrt{2}G_F N_A \sum_{f \in {e,p,n}} g_V^f \nonumber \\
           &= \sqrt{2}G_F N_A\left[ -\frac{1}{2}+2\sin^2{(\theta_W)} + \frac{1}{2}-2\sin^2{(\theta_W)} -\frac{1}{2} \right] \nonumber \\
           &= -\frac{1}  {\sqrt{2}} G_F N_e\,,
\end{align}
where the electrical neutrality condition allows us to simply sum the vectorial couplings together, cancelling the electron and proton contributions (and hence, also the $\theta_W$ dependence).
